---
title: "Data Wrangling"
subtitle: Winter Quarter 2026, Georgia Vasey
format: pdf
gfm: default
pdf: default
execute:
  warning: false
---

## ⍟ Data Wrangling

I have multiple raw datasets that I would eventually like to merge into one or two dataframes to complete my analyses on.

## [Response Variables]{.underline}

## 1 - Soil Carbon

First, I need to read in the libraries to help with data wrangling.

```{r warning=FALSE}
## Load packages from library to help with data wrangling
library("tidyverse")
library(here)
```

Now, I'm bringing in the raw soil carbon dataset from Brookside Labs, OH. I want to remove the top seven rows, which is the metadata from Brookside Labs. I need to mutate the column names so that the rest of the raw datasets can match.

```{r}
# Import the data frames for soil carbon
# I collected data in 2023 and 2024 before and after the burns
soil_C_N_2024 <- read_csv(here("data", "soil_C_N_2024.csv"),
                          skip = 7) # Removes top 7 rows with metadata
soil_C_N_2023 <- read_csv(here("data", "soil_C_N_2023.csv"),
                          skip = 7) # Removes top 7 rows with metadata
```

Now, I want to manipulate the dataframe

```{r}
# 2024 Burns
# Create clean dataframe for soil carbon in 2024 that only includes Carbon
soil_C_2024 <- soil_C_N_2024 %>%
  # Remove everything except for %C
  select(-2, -3, -5, -6) %>%
  # Convert any hyphens to underscores (safe even if there are none)
  ##Create a standardized version of Sample Location so everything uses "_"
  mutate(Sample_Location_std = str_replace_all(`Sample Location`, "-", "_")
  ) %>% 
  # Also need to rename CSR, to remove CSR_SC (I didn't end up sampling at two different sites within CSR)
    mutate(
  Sample_Location_std = if_else(
    #The condition: If a Sample Location starts with CSR_SC_
    str_starts(Sample_Location_std, "CSR_SC_"), #value_if_true
    # If FALSE, replace with the value_if_true (CSR_SC_) with value_if_false (CSR_)
    str_replace(Sample_Location_std, "^CSR_SC_", "CSR_"),
    # If string does not include CSR_SC, return original value unchanged
    Sample_Location_std
  ),
  # Make sure every row ends with a time token (_pre or _post).
     Sample_Location_std = if_else(
       #If it already ends in _pre or _post, keep it.
      str_detect(Sample_Location_std, "_(pre|post)$"),
      Sample_Location_std,
       #Otherwise, for all controls like PL_C_T1_15, append _pre.
      paste0(Sample_Location_std, "_pre")
    )
  ) %>%
  # Separate the standardized 'Sample Location' into individual columnns
  separate(
    col = Sample_Location_std,
    into = c("Site", "Treatment", "Transect", "Sample", "Time"),
    sep = "_",
    extra = "merge",
    fill = "right"
  ) %>%
  #Recode + convert to numeric
  mutate(
    Treatment = recode(Treatment, "B" = "Burn", "C" = "Control"),
    Time      = recode(Time, "pre" = "Pre-burn", "post" = "Post-burn"),
    Sample    = parse_number(Sample),
    `Carbon (%)` = parse_number(`Carbon   (%)`)
  ) %>%
  # 5) Drop the original carbon column + optional keep original Sample Location if you want
  select(Site, Treatment, Transect, Sample, Time, `Carbon (%)`, `Sample Location`)

# Create Transect_ID
soil_C_2024 <- soil_C_2024 %>%
  mutate(
    Transect_ID = paste0(
      Site, "_",
      recode(Treatment, "Burn" = "B", "Control" = "C"), "_",
      Transect
    )
  )

# Getting rid of the two columns with NA
soil_C_2024 <- soil_C_2024 %>%
  filter(
    !is.na(Site),
    !is.na(Treatment),
    !is.na(Transect),
    !is.na(Sample),
    !is.na(Time)
  )

# 2023 Burns
# Create clean dataframe for soil carbon in 2023 that only includes Carbon
soil_C_2023 <- soil_C_N_2023 %>%
  # Only include %C
  select(-1, -7, -8, -10, -11) %>% 
  mutate('Carbon (%)' = as.numeric(`Carbon   (%)`)) %>% 
  # Get rid of the Carbon column that is character
  select(-6)

# Make consistent with soil_C_2024
soil_C_2023 <- soil_C_2023 %>%
  rename(Transect = Transect_ID)

# Create Transect_ID for 2023 Burns
soil_C_2023 <- soil_C_2023 %>%
  mutate(
    Transect_ID = paste0(
      Site, "_",
      recode(Treatment, "Burn" = "B", "Control" = "C"), "_",
      Transect
    )
  )
```

Now it's time to merge 2023 and 2024 burns with bulk density, so that I can calculate soil C Mg ha. The bulk density is required to calculate soil organic carbon stocks in Megagrams of carbon per hectare. Bulk density is the dry weight of a known volume of soil, which is multiplied by the % total carbon. For my soil samples, the volume for 2023 burns was 15 cm depth, whereas the 2024 burns was s 5 cm depth. This is adjusted for in the bulk density calculation.

```{r}
# Merge soil_C_2023 and soil_C_2024
soil_C <- bind_rows(soil_C_2023, soil_C_2024)

# Import bulk density data frame
bulk_density <- read_csv(here("data", "bulk_density.csv"))

# Only keep bulk density
bulk_density_red <- bulk_density %>% 
  # Keep only Site, Treatment, Name, Bulk Density
  select(3, 4, 5, 12) %>% 
  # Rename bulk density
  rename(bulk_density = `Correct bulk density`) %>%
  # Correct Site names for Wilder Ranch to not include underscore
  mutate(
    Site = str_replace_all(Site, "_", "")
  ) %>% 
  # Add a column called Transect_ID
  mutate(
    Transect = str_extract(Name, "T[1-5]")
  ) %>% 
  #Relocate column Transect_ID
  relocate(Transect, .after = Name)

# Create Transect_ID for reduced bulk density
bulk_density_red <- bulk_density_red %>%
  mutate(
    Transect_ID = paste0(
      Site, "_",
      recode(Treatment, "Burn" = "B", "Control" = "C"), "_",
      Transect
    )
  )

# Join bulk density into the dataframe
# Also removed MC (Molino Creek), since I didn't send in samples
soil_C_bulk <- soil_C %>%
  left_join(
    bulk_density_red,
    by = c("Site", "Treatment", "Transect", "Transect_ID")
  ) %>% 
  # Remove Sample Location
  select(-8) 

#Calculate a new column for soil carbon (Megagrams) per hectare 
soil_C_bulk <- soil_C_bulk %>% 
  mutate(C_Mg_per_ha = `Carbon (%)`*bulk_density*10 )
```

Now I want to create another table that evaluates the change in soil C. I want this for both the % total carbon and soil C (Mg) per hectare. I'm taking the difference of Post-burn and Pre-burn.

```{r}
soil_C_change <- soil_C_bulk %>%
  # Select columns to calculate the change in soil C
  select(Site, Treatment, Transect, Transect_ID, Sample, Time, Name, `Carbon (%)`, C_Mg_per_ha) %>%
  # Make one row per Site-Treatment-Transect-Sample, with separate columns for Pre vs Post
  pivot_wider(
    names_from  = Time,
    values_from = c(`Carbon (%)`, C_Mg_per_ha),
    names_sep   = "_"
  ) %>%
  # Compute change as (Post - Pre)  
  mutate(
    # New columns for % C change
    final_Carbon_pct  = `Carbon (%)_Post-burn` - `Carbon (%)_Pre-burn`,
    # New column for soil C (Mg) per ha change
    final_C_Mg_per_ha = `C_Mg_per_ha_Post-burn` - `C_Mg_per_ha_Pre-burn`
  ) %>% 
# Remove Sample Location
  select(-6) %>% 
# Remove all Site = MC  
  dplyr::filter(!is.na(Site), Site != "MC")
```

## 2 - Root Biomass

I'm bringing in the raw dataset for root biomass. I only measured root biomass post-burn for the 2024 burns in both the burn and control. I want to evaluate whether there is an increase in root growth post-fire.

The volume of each soil core (5 cm diameter, 15 cm length) was 294.52 cm\^3 ; V = π r2 h or V = π (2.5 x 2.5)(15). Root biomass per unit of ground area (g/m\^2) from V = A d or V = (100 cm)(100cm)(15 cm) = 150,000 cm\^3. Thus, each core sample represents 294.52 cm\^3 of 150,000 cm\^3. To convert dry roots to g/m\^2, I multiplied dry weight (g) by the inverse of the fraction (150,000/294.52) to get the root biomass of each sample (Roots_g_per_m2).

```{r}
#Read in data
roots <- read_csv(here("data", "root_biomass.csv"))
```

## 3 - Aboveground Biomass

I'm bringing in the raw dataset for aboveground biomass, also called residual dry matter (RDM). I collected RDM pre-burn in 2023 and 2024. I then collected RDM one year later for 2024 burns and 2 years later for 2023 burns in the fall. Unfortunately, I forgot to sample at the 10- , 20-, 30-, and 40-m for the 2023 burns, so I had to remove this data. Now I want to estimate the change in aboveground biomass (2024) for the burn versus control sites.

```{r}
#Read in data
aboveground_measurements <- read_csv(here("data", "rdm_fireintensity_soilmoisture.csv"))

head(aboveground_measurements)

#Create a new dataframe for RDM change only 
RDM_change <- aboveground_measurements %>%
  # Select columns to calculate the change in RDM
  select(Site, Treatment, Transect, Transect_ID, Sample, RDM_g_preburn, RDM_g_postburn) %>%
  # Compute change as (Post - Pre)  
  mutate(
    # New columns for the change in RDM (final_RDM)
    final_RDM  = RDM_g_postburn - RDM_g_preburn)
```

## [Predictor Variables]{.underline}

## 1 - Weather

I found the burn date for each site from the CAL FIRE FRAP (Fire and Resource Assessment Program) database. I then obtained local weather station data:

-   RAWS for CSR (Pulgas California; 37.475, -122.298056), CF (Coast Dairies California; 37.030833, -122.187778), and AM (Big Sur California; 36.235556, -121.785)

<!-- -->

-   CIMIS for WR22, WR13, WR5, and MF (De Laveagea (104); 36.997444, -121.99676), PL (Pacific Grove (193); 36.63322, -121.93486), and TKR (Pescadero (253); 37.255333, -122.3708)

I have weather data for the day of the burn, number of days since rainfall event ≥ 7 mm (antecedent precip), the 5-day average for weather variables (including the burn date) during the week leading up to the burn, and finally the 5-year average for all weather variables.

```{r}
# Load libraries to deal with excel spreadsheet
library(readxl)
library(dplyr)
library(lubridate)
library(stringr)

#Read in data
file_path <- here::here("data", "Weather.xlsx")
 # Need a string for later, not a data frame

# Create Burn Date (Site and Day of Burn)
 # Burn Date was obtained through CAL FIRE FRAP Database
burn_dates <- tibble::tribble(
  ~Site,  ~Date,
  "CSR",  "10/11/2024",
  "TKR",  "10/25/2024",
  "CF",   "11/01/2023",
  "WR5",  "11/13/2024",
  "MF",   "11/20/2024",
  "WR22", "11/09/2023",
  "WR13", "11/13/2024",
  "PL",   "11/19/2024",
  "AM",   "12/13/2023"
) %>%
  mutate(
    Date = mdy(Date),
    Year = year(Date)
  )

# Obtain weather data from the Site_daily sheets
 # WR_23_daily is WR22
 # WR_MF_24_daily is WR5, MF, WR13
site_sheet_map <- tibble::tribble(
  ~Site,  ~daily_sheet,
  "CSR",  "CSR_daily",
  "TKR",  "TKR_daily",
  "CF",   "CF_daily",
  "WR5",  "WR_MF_24_daily",
  "MF",   "WR_MF_24_daily",
  "WR13", "WR_MF_24_daily",
  "WR22", "WR_23_daily",
  "PL",   "PL_daily",
  "AM",   "AM_daily"
)

# Develop a function: read a daily sheet safely
read_daily_sheet <- function(sheet_name) {
  readxl::read_excel(file_path, sheet = sheet_name) %>%
    # standardize column names a bit (optional)
    rename_with(~ str_replace_all(.x, "\\s+", "_"))
}

# Develop a function: pull burn-day row for one site
get_burn_day_values <- function(site, burn_date, sheet_name) {

  daily <- read_daily_sheet(sheet_name) %>%
    mutate(date = as.Date(date))  # ensure Date class

  # Columns needed from each daily sheet (plus date for filtering)
  keep_cols <- c(
    "date", "Jul", "wind_speed",
    "mean_temp", "max_temp", "min_temp",
    "mean_humidity", "max_humidity", "min_humidity",
    "precipitation"
  )

  # Some sheets include extra columns (e.g., stn_id, year, Unnamed:*).
  # We only select what we need IF they exist.
  burn_row <- daily %>%
    filter(date == burn_date) %>% # Burn Date
    select(any_of(keep_cols)) # Selected weather data

  # If missing, return an NA row with the expected columns to prevent bind_rows errors
  if (nrow(burn_row) == 0) {
    burn_row <- tibble(
      date = burn_date,
      Jul = NA_real_,
      wind_speed = NA_real_,
      mean_temp = NA_real_,
      max_temp = NA_real_,
      min_temp = NA_real_,
      mean_humidity = NA_real_,
      max_humidity = NA_real_,
      min_humidity = NA_real_,
      precipitation = NA_real_
    )
  }

  burn_row %>%
    mutate(Site = site) %>%
    select(Site, everything())
}

# Function for antecedent precipitation and days since last precipitation event
  # days_since_precip = days between burn date and most recent prior date with precip>=7
 # antec_precip      = that precipitation amount
get_antec_precip <- function(site, burn_date, sheet_name, threshold = 7) {

  daily <- read_daily_sheet(sheet_name) %>% #Get the daily sheet
    mutate(date = as.Date(date)) #Make sure that Date is taken from date
  # Look for the prior rain event 
  prior_event <- daily %>%
    filter(date < burn_date) %>%
    # Threshold is >= to 7 mm precipitation event
    filter(!is.na(precipitation) & precipitation >= threshold) %>% 
    arrange(desc(date)) %>%
    slice(1)
  # Deal with if there is no prior rain event 
  if (nrow(prior_event) == 0) {
    return(tibble(
      Site = site,
      days_since_precip = NA_integer_,
      antec_precip = NA_real_
    ))
  }

  tibble(
    Site = site,
    # Gives number of days since burn date with precipitation >= 7 mm
    days_since_precip = as.integer(burn_date - prior_event$date), 
    # Records the value of the event >= 7mm
    antec_precip = as.numeric(prior_event$precipitation)
  )
}

# Finally build the burn-day daily-weather table (Jul -> precipitation)
burn_day_weather <- burn_dates %>% #Original data frame with Site and Burn Date
  # Left two columns are Site and Date for burn date
  left_join(site_sheet_map, by = "Site") %>%
  rowwise() %>%
  do({
    site <- .$Site #Site
    burn_date <- .$Date #Burn Date
    sheet_name <- .$daily_sheet #Values from Site_daily sheets

    out <- get_burn_day_values(site, burn_date, sheet_name)
    # "date", "Jul", "wind_speed", "mean_temp", "max_temp", "min_temp", "mean_humidity", "max_humidity", "min_humidity","precipitation"
    out
  }) %>%
  ungroup() %>%
  rename(Date = date)  # burn date column name as requested

# Now add days_since_precip + antec_precip to original data frame (burn dates)
 # Will eventually merge all of the tables 
antec_tbl <- burn_dates %>% #Again, the Site and Burn Date data frame
  left_join(site_sheet_map, by = "Site") %>% #Join by Site
  rowwise() %>%
  do({
    get_antec_precip(.$Site, .$Date, .$daily_sheet, threshold = 7)
  }) %>%
  ungroup()

# Last rows to be added are the 5-day means from weather_5_days (cols 3:9)
 # Filters each site to the 5-day window ending on burn date
 # More accurate representation of the weather leading up to the burn date
weather_5_days <- readxl::read_excel(file_path, sheet = "weather_5_days") %>%
  mutate(date = as.Date(date)) # Get the sheet and look for the burn date
#Create the 5 day weather summary table
five_day_summary <- burn_dates %>% #Again, to merge with original Site and Burn Date table
  rowwise() %>%
  do({
    site <- .$Site
    burn_date <- .$Date

    w <- weather_5_days %>% # w is which days to pull weather data from 
      filter(Site == site) %>%
      # Look at the Burn Date and the 4 prior days (5 day average)
      filter(date >= (burn_date - days(4)) & date <= burn_date)

    # If fewer than 1 row exists, return NA columns
    if (nrow(w) == 0) {
      return(tibble(
        Site = site,
        `5d_mean_temp` = NA_real_,
        `5d_max_temp` = NA_real_,
        `5d_min_temp` = NA_real_,
        `5d_mean_humidity` = NA_real_,
        `5d_max_humidity` = NA_real_,
        `5d_min_humidity` = NA_real_,
        `5d_precipitation` = NA_real_
      ))
    }

    # Columns 3:9 are:
    # mean_temp_5, max_temp_5, min_temp_5, mean_humidity_5, max_humidity_5, min_humidity_5, precipitation_5
    s <- w %>%
      # Take the mean of the 5 values from each of these columns
      summarise(
        mean_temp_5 = mean(mean_temp_5, na.rm = TRUE),
        max_temp_5 = mean(max_temp_5, na.rm = TRUE),
        min_temp_5 = mean(min_temp_5, na.rm = TRUE),
        mean_humidity_5 = mean(mean_humidity_5, na.rm = TRUE),
        max_humidity_5 = mean(max_humidity_5, na.rm = TRUE),
        min_humidity_5 = mean(min_humidity_5, na.rm = TRUE),
        precipitation_5 = mean(precipitation_5, na.rm = TRUE)
      ) %>%
      # Rename them, since I will also create 5 year averages in weather eventually
      rename(
        `5d_mean_temp` = mean_temp_5,
        `5d_max_temp` = max_temp_5,
        `5d_min_temp` = min_temp_5,
        `5d_mean_humidity` = mean_humidity_5,
        `5d_max_humidity` = max_humidity_5,
        `5d_min_humidity` = min_humidity_5,
        `5d_precipitation` = precipitation_5
      ) %>%
      mutate(Site = site)

    s #Return the values
  }) %>%
  ungroup()

#### Final table that combines original data frame, burn date weather, antecedent precip, and 5 day weather averages ####
weather <- burn_dates %>%
  select(Site, Date, Year) %>% #Select all three columns from burn_dates
  # Join with burn day weather
  left_join(burn_day_weather, by = c("Site", "Date")) %>%
  # Join with antecedent precip table
  left_join(antec_tbl, by = "Site") %>%
  # Join with 5 day average for weather
  left_join(five_day_summary, by = "Site") %>%
  # Make sure all of the columns exist
  select(
    Site, Date, Year, Jul,
    wind_speed,
    mean_temp, max_temp, min_temp,
    mean_humidity, max_humidity, min_humidity,
    precipitation,
    days_since_precip, antec_precip,
    `5d_mean_temp`, `5d_max_temp`, `5d_min_temp`,
    `5d_mean_humidity`, `5d_max_humidity`, `5d_min_humidity`,
    `5d_precipitation`
  )

# Lastly, I need to obtain the 5-year monthly weather average, and add it to the weather data frame

# Read in data
file_5y <- here::here("data", "10_5_Weather.xlsx")

# Need to specify the values from the columns that I want to average
vars_5y <- c(
  "mean_temp", "max_temp", "min_temp",
  "mean_humidity", "max_humidity", "min_humidity",
  "precipitation"
)

# Develop a function that reads the sheet, and computes means of the selected columns
 # This also ignores all other columns that are not from vars_5y
compute_5y_means <- function(sheet_name) {
  df <- readxl::read_excel(file_5y, sheet = sheet_name) %>%
    rename_with(~ stringr::str_replace_all(.x, "\\s+", "_"))

  # Compute means for the target variables
  out <- df %>%
    summarise(across(
      any_of(vars_5y),
      ~ mean(as.numeric(.x), na.rm = TRUE)
    ))

  # Rename to 5y_* format
  out %>%
    # Rename the variables with 5y_ in front to not confuse with 5d_ (5 days)
    rename_with(~ paste0("5y_", .x), .cols = everything()) %>%
    mutate(source_sheet = sheet_name) %>%
    select(source_sheet, everything())
}

# Compute 5-year means for each source sheet 
means_pulgas   <- compute_5y_means("Pulgas_5")      # CSR
means_ccd      <- compute_5y_means("CCD_3")         # CF
 # Note, that CCD only went back 3 years 
means_bigsur   <- compute_5y_means("Big Sur_5")     # AM
means_pesc     <- compute_5y_means("Pescadero_5")   # TKR
means_dela     <- compute_5y_means("DeLa_5")        # WR22, WR13, WR5, MF
means_pacgrove <- compute_5y_means("PacGrove_5")    # PL

#Making sure to specify which sites are associated with the 5y source sheets
five_year_by_site <- bind_rows(
  means_pulgas   %>% mutate(Site = "CSR"),
  means_ccd      %>% mutate(Site = "CF"),
  means_bigsur   %>% mutate(Site = "AM"),
  means_pesc     %>% mutate(Site = "TKR"),
  means_pacgrove %>% mutate(Site = "PL"),

  # DeLa_5 applies to multiple sites:
  means_dela %>% mutate(Site = "WR22"),
  means_dela %>% mutate(Site = "WR13"),
  means_dela %>% mutate(Site = "WR5"),
  means_dela %>% mutate(Site = "MF")
) %>%
  select(
    Site,
    starts_with("5y_")
  )

#### Final, final weather table that includes 5-year normals for weather ####
weather_final <- weather %>%
  left_join(five_year_by_site, by = "Site")

# Double check that it merged corectly for all 9 sites
print(weather_final %>% select(Site, starts_with("5y_")) %>% distinct())

##Note - there is no Transect_ID for this predictor variable
 # Additionally, there is no Treatment 
```

## 2 - Topography

I created polygons around each burn and control prairie in ArcGIS Pro. I then downloaded topography raster (DEM) from The National Map (3DEP) with 10m resolution (10 m x 10 m = 100 m2 on the ground): <https://index.nationalmap.gov/arcgis/rest/services/3DEPElevationIndex/MapServer> (includes elevation, slope, and aspect). I clipped the raster to the prairie perimeters, and used Zonal Statistics to get several points for each DEM variable. In this code, I'm taking the mean of elevation, slope, and aspect.

```{r}
# Load library
library(readxl)

#Read in the data
dem_path <- here::here("data", "DEM_10m.xlsx") # Needs to be a string

# Create a function to standardize the Site names
clean_site <- function(x) {
  x %>%
    #Make column text
    as.character() %>%
    #Remove leading/trailing spaces 
    str_trim() %>%
    # Remove the underscores in WR
    str_replace_all("^WR_5$",  "WR5") %>%
    str_replace_all("^WR_13$", "WR13") %>%
    str_replace_all("^WR_22$", "WR22")
}

# Create a function to summarize a DEM sheet by Site and Treatment
 # Detect variable column (Elevation, Slope, and Aspect)
 # Calculates mean per Site
 # Sheets ending in *_unburn are Treatment = Control
 #Sheets ending in *_burn are Treatment = Burn
process_dem_sheet <- function(sheet_name) {

  # Keep the original sheet name EXACTLY for read_excel()
   # Had trailing spaces
  sheet_name_orig <- sheet_name

  # Use a trimmed/lowercase copy ONLY for pattern matching
  sheet_low <- tolower(stringr::str_trim(sheet_name_orig))

  # Read one sheet (must match exactly, including trailing spaces)
  df <- readxl::read_excel(dem_path, sheet = sheet_name_orig)

  # Remove OBJECTID if it exists
  if ("OBJECTID" %in% names(df)) {
    df <- df %>% dplyr::select(-OBJECTID)
  }

  #Standardize Site names + remove MC
  df <- df %>%
    dplyr::mutate(Site = clean_site(Site)) %>%
    dplyr::filter(!is.na(Site), Site != "MC")

  # Infer Treatment from sheet name 
  Treatment_value <- dplyr::case_when(
    stringr::str_detect(sheet_low, "unburn") ~ "Control",
    stringr::str_detect(sheet_low, "burn")   ~ "Burn",
    TRUE ~ NA_character_
  )
  #Safety check
  if (is.na(Treatment_value)) {
    stop(paste0(
      "Could not infer Treatment for sheet '", sheet_name_orig, "'. ",
      "Expected sheet name to include 'unburn' or 'burn'."
    ))
  }

  df <- df %>% dplyr::mutate(Treatment = Treatment_value)

  # Identify the variable column (besides Site and Treatment)
  variable_col <- setdiff(names(df), c("Site", "Treatment"))
  
  #Safety check
  if (length(variable_col) != 1) {
    stop(paste0(
      "Sheet '", sheet_name_orig, "' has unexpected columns besides Site/Treatment: ",
      paste(variable_col, collapse = ", "),
      ". Expected exactly ONE variable column."
    ))
  }

  # Summarise mean by Site + Treatment (integer)
  df %>%
    dplyr::mutate(value = as.numeric(.data[[variable_col]])) %>%
    dplyr::group_by(Site, Treatment) %>%
    dplyr::summarise(mean_value = round(mean(value, na.rm = TRUE), 0), .groups = "drop") %>%
    dplyr::mutate(variable = tolower(variable_col)) %>%
    dplyr::select(Site, Treatment, variable, mean_value)
}

#Get list of sheets in the excel workbook
sheet_names <- excel_sheets(dem_path)

# Process the sheets
dem_long <- purrr::map_dfr(sheet_names, process_dem_sheet)

# Convert to wide format
DEM <- dem_long %>%
  pivot_wider(
    names_from = variable,
    values_from = mean_value
  ) %>%
  select(Site, Treatment, any_of(c("elevation", "slope", "aspect"))) %>%
  arrange(Site, Treatment)

## Note - there is no Transect_ID for this predictor variable
```

## 3 - Soil Properties

First, I want to summarize the soil texture data which was measured in the lab using the hydrometer method. I have to clean up the dataset to include Site, Treatment, Transect_ID, and then only obtain the % Sand, % Silt, and % Clay numbers.

```{r}
#Read in data
soil_texture <- read_csv(here("data", "soil_texture.csv"))

# Standardize column names
 # This keeps the original names, but makes it easier to reference if there are extra spaces.
names(soil_texture) <- str_trim(names(soil_texture))

texture <- soil_texture %>% 
# Remove excluded transects
  filter(
    !str_detect(Transect, "^(CCD|QV|SLP|SPR)"), # Get rid of these
    # Get rid of the other Andrew Molera site (Plot 4)
    !(str_detect(Transect, "^AM") & str_detect(tolower(Transect), "(p4|plot\\s*4)"))
  ) %>%
#Remove the "_SC" token so Treatment/Transect tokens align
  mutate(
    Transect_clean = Transect %>%
      str_replace_all("[_\\s-]+SC(?=[_\\s-]+)", "_") %>%  # SC in the middle
      str_replace_all("[_\\s-]+SC$", "") %>%             # SC at end
      str_replace_all("^SC[_\\s-]+", "")                 # SC at beginning (rare)
  ) %>%

# Tokenize & pull out key parts
  mutate(
    .tokens = str_split(Transect_clean, pattern = "[_\\s-]+"),

    # Base site token (first token)
    Site_raw = toupper(map_chr(.tokens, ~ .x[1] %||% NA_character_)),

    # Treatment token should now be second token: B or C (Burn/Control)
    Trt_code = toupper(map_chr(.tokens, ~ .x[2] %||% NA_character_)),

    Treatment = case_when(
      Trt_code == "B" ~ "Burn",
      Trt_code == "C" ~ "Control",
      TRUE ~ NA_character_
    ),

    # Transect token should now be third token: T1..T5
    Transect_only = toupper(map_chr(.tokens, ~ .x[3] %||% NA_character_)),

    # Extract plot number (for WR only), supports:
    # P13 / P 13 / Plot 13 / P22 / 3/5 -> 5
    Plot_raw = str_to_lower(Transect_clean),
    Plot = case_when(
      str_detect(Plot_raw, "3\\s*/\\s*5") ~ "5", #Change all 3 to 5
      str_detect(Plot_raw, "\\bp\\s*\\d+\\b") ~ str_match(Plot_raw, "\\bp\\s*(\\d+)\\b")[,2],
      str_detect(Plot_raw, "plot\\s*\\d+") ~ str_match(Plot_raw, "plot\\s*(\\d+)")[,2],
      TRUE ~ NA_character_
    ),

    # Normalize non-WR site codes (These sites have 3 letters)
    Site_raw = case_when(
      Site_raw == "CS" ~ "CSR",
      Site_raw == "TK" ~ "TKR",
      TRUE ~ Site_raw
    ),

#Build Transect_ID and Site
 #WR: Transect_ID = WR_B_T1_13 ; Site = WR13
 #Others: Transect_ID = CSR_B_T1 ; Site = CSR, etc.
    Transect_ID = case_when(
      Site_raw == "WR" & !is.na(Plot) ~ str_c("WR", Trt_code, Transect_only, Plot, sep = "_"), #Create the site name for WR to be WR5, WR13, and WR22
      TRUE ~ str_c(Site_raw, Trt_code, Transect_only, sep = "_")
    ),

    Site = case_when(
      Site_raw == "WR" & !is.na(Plot) ~ paste0("WR", Plot),
      TRUE ~ Site_raw
    )
  ) %>%
# Keep only rows that successfully parsed
  filter(
    !is.na(Site),
    !is.na(Treatment),
    str_detect(Transect_only, "^T[1-5]$"),
    Trt_code %in% c("B", "C")
  ) %>%
# Select final columns and rename texture vars
  transmute(
    Site,
    Treatment,
    Transect_ID,
    sand = .data[["% Sand"]],
    clay = .data[["% Clay"]],
    silt = .data[["% Silt"]]
  ) %>%
  arrange(Site, Treatment, Transect_ID)

# Remove excluded transects
 # Remove any Transect starting with CCD, QV, SLP, SPR
  #Remove AM Plot 4
tex2 <- soil_texture %>%
  filter(
    !str_detect(Transect, "^(CCD|QV|SLP|SPR)"),
    !(str_detect(Transect, "^AM") & str_detect(tolower(Transect), "(P4|plot\\s*4)"))
  )

# Helper function for %||% (base-style fallback for NULL)
 # Return a if it exists and has length > 0.
 # Inside str_split() results, some rows might not have enough tokens.
`%||%` <- function(a, b) if (!is.null(a) && length(a) > 0) a else b

# Need to first clean up Transect_ID in texture to match nutrients (remove _Plot)
texture <- texture %>%
mutate(
    # If it's WR_*_*_<plot>, rewrite as WR<plot>_*_*
    Transect_ID = case_when(
      # ([BC]) = treatment code ; (T[1-5]) = Transect ; (\\d+) = plot number
      str_detect(Transect_ID, "^WR_[BC]_T[1-5]_\\d+$") ~ 
        str_replace(Transect_ID, "^WR_([BC])_(T[1-5])_(\\d+)$", "WR\\3_\\1_\\2"),
      TRUE ~ Transect_ID
    ),
    # Make Site match the new WR<plot> prefix when applicable
    Site = case_when(
      str_detect(Transect_ID, "^WR\\d+_") ~ str_extract(Transect_ID, "^WR\\d+"),
      TRUE ~ Site
    )
  )

```

I now want to link this with the baseline soil nutrients data. This data was obtained from additional soil samples pre-burn, which was sent to Brookside Labs, OH. I also have post-burn soil nutrients for the **BURN**-only transects. I could possibly look at the change in soil nutrients, but most likely I should just use the pre-burn soil nutrients for Treatments = Burn and Control for all transects.

```{r}
#Read in data
soil_nutrients <- read_csv(here("data", "soil_nutrients.csv"),
                          skip = 7) # Removes top 7 rows with metadata

# Trim any header whitespace
names(soil_nutrients) <- str_trim(names(soil_nutrients))

# Remove rows without Sample Location (the last two)
soil_nutrients <- soil_nutrients %>%
  filter(!is.na(`Sample Location`), str_trim(`Sample Location`) != "")

# Clean Sample Location strings
 # Remove CSR "_SC" 
soil_nutrients  <- soil_nutrients  %>%
  mutate(
    sample_loc_raw = str_trim(`Sample Location`),
    sample_loc = str_replace_all(sample_loc_raw, "_SC_", "_")  # only affects CSR_SC_...
  )

# Remove Sample Description and Lab Number
soil_nutrients <- soil_nutrients %>%
  select(
    -`Sample Description #1`,
    -`Lab Number`
  )

# Create Site, Treatment, Sample, Time
soil_parsed <- soil_nutrients %>% 
    mutate(
  # Site: everything before the first underscore
    Site = str_extract(sample_loc, "^[^_]+"),

    # Treatment code extraction
     # Burn if B, Control if C
     # If "_C_B" appears, use the first (C) => Control
    # Implementation:
     # Prefer the first occurrence of "_B" or "_C" as a token
    Trt_code = str_match(sample_loc, "_([BC])(?=_)")[, 2],

    Treatment = case_when(
      Trt_code == "B" ~ "Burn",
      Trt_code == "C" ~ "Control",
      TRUE ~ NA_character_
    ),
    
    # Transect: find _T1.._T5 anywhere in the string
    Transect = str_match(sample_loc, "_(T[1-5])(?:_|$)")[, 2],

    # Transect_ID: Site_B/C_T# ----
    Transect_ID = str_c(Site, Trt_code, Transect, sep = "_"),


    # Sample:
     # 1 if contains "_1to4"
     # Otherwise 15/30/45 if contains "_15" or "_30" or "_45"
    Sample = case_when(
      str_detect(sample_loc, "_1to4") ~ 1L,
      str_detect(sample_loc, "_15(?!\\d)") ~ 15L,
      str_detect(sample_loc, "_30(?!\\d)") ~ 30L,
      str_detect(sample_loc, "_45(?!\\d)") ~ 45L,
      TRUE ~ NA_integer_
    ),

    # Time:
     # Strings ending with "_post" for Post-burn.
     # Best-practice rule: if it ends in "_post" => Post-burn, else Pre-burn
    Time = case_when(
      str_detect(sample_loc, "_post$") ~ "Post-burn",
      TRUE ~ "Pre-burn"
    )
  )

# Getting rid of helper columns
nutrients <- soil_parsed %>%
  # Keep all original columns, plus new ones
  select(Site, Treatment, Transect, Transect_ID, Sample, Time, everything()) %>%
  # Drop helper columns
  select(-`Sample Location`, -sample_loc_raw, -sample_loc, -Trt_code)

# Clean up nutrients mistakes 
nutrients <- nutrients %>%
# Replace Site WR with WR22
  mutate(
    Site = if_else(Site == "WR", "WR22", Site)
  ) %>%
# Update Transect_ID to include WR22 instead of WR
  mutate(
    Transect_ID = stringr::str_replace(Transect_ID, "^WR_", "WR22_")
  ) %>%
# Remove MC rows
  filter(Site != "MC")

```

Now, I want to combine soil texture and soil nutrients, to get one table for soil properties.

```{r}
# Join nutrients with soil texture 
soil_properties <- nutrients %>%
  left_join(
    texture,
    by = c("Site", "Treatment", "Transect_ID")
  )

# Double-check output
print(soil_properties)
```

## 4 - Composite Burn Index

My final predictor variable needs to quantify the various burns. It was recommended that I compile multiple components into a Composite Burn Index. I measured pre- and post litter right before and after the burn for both Burn and Control. I also have pre-burn residual dry matter, as well as post-burn burn severity in cover classes (% Black, % Living Green, % New Green, % Bare Ground).

To quantify fire effects across vegetation, litter, and fuel characteristics, I developed a Composite Burn Index (CBI) using a scaled continuous scoring system. For each component (e.g., residual dry matter, litter consumption, and burn severity), raw field measurements were rescaled from 0 to 3 using min-max normalization, such that:

-   Score= 3 × \[Value−min(Value)/\[max(Value)−min(Value)​\]

    -    Minimum and maximum values are the minimum and maximum score from all of the transects combined

These scale each fuel/severity component based on assumptions/rationale:

-   **RDM_Score**: Higher biomass = higher potential intensity.

-   **Litter_Score**: Greater litter reduction = more intense burn.

-   **Severity_Score**: Combines Black + Bare_Ground − (Living + New Green + Dried Unburned).

    -   Higher = more visible burn effects.

According to the *NPS Fire Monitoring Handbook*, these are the standard on how to assess burn severity in grasslands:

-   **Unburned:** [VEGETATION]{.underline}: not burned; [SUBSTRATE]{.underline}: not burned

-   **Scorched:** [VEGETATION]{.underline}: Foliage is scorched; litter partially blackened; [SUBSTRATE]{.underline}: duff nearly unchanged; leaf structures unchanged

-   **Lightly Burned**: [VEGETATION]{.underline}: Grasses with approximately two inches of stubble; foliage and smaller twigs of associated species partially to completely consumed; some plant parts may still be standing; bases of plants are not deeply burned and are still recognizable; [SUBSTRATE]{.underline}: Litter charred to partially consumed, but some plant parts are still discernible; charring may extend slightly into soil surface, but soil is not visibly altered; surface appears black (this soon becomes inconspicuous); burns may be spotty to uniform depending on the grass continuity.

-   **Moderately Burned**: [VEGETATION]{.underline}: Unburned grass stubble, usually less than 2 inches tall, and mostly confined to an outer ring; for other species, foliage completely consumed, plant bases are burned to ground level and obscured in ash immediately after burning; burns tend to be uniform; [SUBSTRATE]{.underline}: leaf litter consumed, leaving coarse, light gray or white colored ash immediately after the burn; ash soon disappears leaving bare mineral soil; charring may extend slightly into soil surface

-   **Heavily Burned**: [VEGETATION]{.underline}: No unburned grasses above the root crown; for other species, all plant parts consumed leaving some or no major stems or trunks, any left are deeply charred; this severity class is uncommon due to the short burnout time of grasses; [SUBSTRATE]{.underline}: leaf litter completely consumed, leaving a fluffy white ash, this soon disappears leaving bare mineral soil; charring extends to a depth of 1 cm (0.5 in) into the soil; this severity class is usually limited to situations where heavy fuel load on mesic sites has burned under dry conditions and low wind.

To interpret the CBI index, I will base scores off of above descriptions, averaging the three scores:

-   0.0 – 0.9: Unburned or lightly scorched

-   1.0 – 1.9: Low to moderate severity

-   2.0 – 2.9: Moderate to high severity

-   3.0: Extreme severity (rare in grasslands)

```{r}
# Load data
burn_severity <- read_csv(here("data", "master_burn_severity.csv"))
litter <- read_csv(here("data", "master_litter_veght.csv"))
# Use rdm from aboveground_measurements (from 3- Aboveground Biomass)
# aboveground_measurements <- read_csv(here("data", "rdm_fireintensity_soilmoisture.csv"))

## Burn Severity Scores (Percent Cover)
# Step 1: # Convert location columns to numeric
burn_severity <- burn_severity %>%
  mutate(across(as.character(seq(5, 50, by = 5)), ~ as.numeric(as.character(.))))

# Now pivot longer
burn_data_long <- burn_severity %>%
  pivot_longer(
    cols = all_of(as.character(seq(5, 50, by = 5))),
    names_to = "Location_m",
    values_to = "Percent_Cover"
  ) %>%
  mutate(Location_m = as.numeric(Location_m)  # Ensure Location is numeric
) %>%
  select(Year, Site, Transect_ID, Location_m, Cover_Class, Percent_Cover)  # Final column order

# Summarize mean by Year, Site, Transect_ID, and Cover_Class
burn_severity_summary <- burn_data_long %>%
  group_by(Year, Site, Transect_ID, Cover_Class) %>%
  summarise(
    Mean_Cover = round(mean(Percent_Cover, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Summarize Severity_Raw and Severity_Score by group
burn_severity_summary <- burn_severity_summary %>%
  #Cover classes needed for the Burn Severity Calculation
  filter(Cover_Class %in% c("Black", "Bare_Ground", "Dried_Unburned", "Living_Green", "New_Green")) %>% 
   #Group by Year, Site, Transect_ID and extract Mean_Cover for required Cover_Class
  select(Year, Site, Transect_ID, Cover_Class, Mean_Cover) %>%                     
  pivot_wider(names_from = Cover_Class, values_from = Mean_Cover, values_fill = 0) %>%
  #How to calculate the score for burn severity with cover class
  mutate(
    Severity_Raw = Black + Bare_Ground - (Dried_Unburned + Living_Green + New_Green)   
  ) %>%
  mutate(
    #Need to rescale for 0-3, since that's all the other scales
    Severity_Score = 3 * (Severity_Raw - min(Severity_Raw, na.rm = TRUE)) /     
      (max(Severity_Raw, na.rm = TRUE) - min(Severity_Raw, na.rm = TRUE))       
    # min is the lowest Severity_Raw across all transects
    # max is the highest Severity_Raw across all transects
  )

# Add Treatment = "Burn" to the burn severity summary to allow merge
burn_severity_summary <- burn_severity_summary %>% mutate(Treatment = "Burn")

## Litter Scores
# Convert Litter_cm to numeric if needed
litter <- litter %>%
  mutate(
    Litter_cm = as.numeric(Litter_cm),
  )

# Summarize Litter (Before/After) for each Year/Site/Treatment/Transect_ID
litter_summary <- litter %>%
  group_by(Year, Site, Treatment, Transect_ID, Timepoint) %>%
  summarise(
    Litter = mean(Litter_cm, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = Timepoint,
    values_from = c(Litter),
    names_glue = "Litter_{Timepoint}"
  )

# Calculate Litter Consumption
litter_summary <- litter_summary %>%
  mutate(
    Litter_Consumption = Litter_Before - Litter_After)

#Create scaled continuous scores based on linear rescaling methods
# This normalized the values between 0 to 1, and then scales them up to 3
# This accounts for the actual spread of values, rather than fixed bins

# Calculate the min and max for the scaling range
litter_min <- min(litter_summary$Litter_Consumption, na.rm = TRUE)
litter_max <- max(litter_summary$Litter_Consumption, na.rm = TRUE)

# Apply continuous 0–3 scaled scores for Litter and Veg Scores
litter_summary <- litter_summary %>%
  mutate(
    Litter_Score = round(3 * (Litter_Consumption - litter_min) / (litter_max - litter_min), 2)
  )

# Residual Dry Matter (RDM) Scores
# Calculate mean for RDM_g
rdm_summary <- aboveground_measurements %>%
  group_by(Year, Site, Transect_ID, Treatment) %>%
  summarise(
    RDM = round(mean(RDM_g_preburn, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Calculate the min and max for the scaling range
rdm_min <- min(rdm_summary$RDM, na.rm = TRUE)
rdm_max <- max(rdm_summary$RDM, na.rm = TRUE)

# Apply continuous 0–3 RDM_Score
rdm_summary <- rdm_summary %>%
  mutate(
    RDM_Score = round(3 * RDM - rdm_min) / (rdm_max - rdm_min), 2)

#Get rid of random column called "2"
rdm_summary <- rdm_summary %>% select(-`2`)

## Combine All Scores
# Merge all datasets by Year, Site, Transect_ID, and Treatment
CBI <- litter_summary %>%
  full_join(rdm_summary, by = c("Year", "Site", "Transect_ID", "Treatment")) %>%
  full_join(burn_severity_summary, by = c("Year", "Site", "Transect_ID", "Treatment"))

# Calculate the CBI as the mean of the four component scores
CBI <- CBI %>%
  mutate(CBI = ifelse(Treatment == "Burn",
                      round((Litter_Score + RDM_Score + Severity_Score) / 4, 2),
                      NA))

# Create categorical column to described Composite Burn Index (CBI) and keep only Burn Transects
CBI <- CBI %>%
  mutate(
    CBI_Category = case_when(
      Treatment == "Control" ~ NA_character_,
      CBI >= 3               ~ "Extreme",
      CBI >= 2               ~ "Moderate_to_High",
      CBI >= 1               ~ "Low_to_Moderate",
      CBI >= 0               ~ "Unburned_to_Scorched",
      TRUE                   ~ NA_character_
    )
  ) %>% 
  # Keep only burned transects
  filter(Treatment == "Burn")

#Look at the table
print(CBI)

## Also a good idea to have pre-burn litter as a predictor variable
head(litter)

#Create a new dataframe for Litter pre-burn
Litter_Pre_burn <- litter %>%
  # Keep only pre-burn measurements
  filter(Timepoint == "Before") %>%
  # Select relevant columns
  select(Year, Site, Treatment, Transect_ID, Location_m, Litter_cm) %>%
  # Average the 10 measurements within each transect
  group_by(Year, Site, Treatment, Transect_ID) %>%
  
  summarise(
    Litter_Pre_burn = mean(Litter_cm, na.rm = TRUE),
    .groups = "drop"
  )

```

## [Analysis Ready Data Frames]{.underline}

I now want to build a transect-level master table (Site x Treatment x Transect_ID), then left-join predictors at the appropriate levels:

-   **Site-level predictors** (weather_final) join by `Site`

-   **Site×Treatment predictors** (DEM) join by `Site + Treatment`

-    **Transect-level predictors** (soil_properties) join by `Site + Treatment + Transect_ID`

    -   Most likely only looking at Time == Pre-burn (Baseline soil conditions)

<!-- -->

-    **Burn-only predictors** (CBI) join by `Year + Site + Transect_ID`

Then, I'll attach each response:

-   `soil_C_change` (all sites/years, both treatments) join by `Site + Treatment + Transect_ID`

<!-- -->

-    `roots` and `RDM_change` (2024 subset) join by `Site + Treatment + Transect_ID`

The script produces two final data frames:

1.  `predictors_master` (baseline predictors at transect-level)
2.  `analysis_carbon` (master + all response columns)

```{r}
##Build analysis-ready dataframe for exploratory plots + models

# Load libraries
library(dplyr)
library(stringr)

### First, need to make some fixes among tables
# (1) Fix the mislabeling of transects. In one of my data frames, the Control Transect_ID is labeled with _C_B, rather than just _C
fix_transect_id <- function(df) {
  if (!("Transect_ID" %in% names(df))) return(df)

  df %>%
    mutate(
      Transect_ID = str_replace(Transect_ID, "_C_B_", "_C_"))
}
# keys
soil_C_change   <- fix_transect_id(soil_C_change)
roots           <- fix_transect_id(roots)
RDM_change      <- fix_transect_id(RDM_change)
soil_properties <- fix_transect_id(soil_properties)
CBI             <- fix_transect_id(CBI)
# DEM and weather_final don't use Transect_ID, so no need

## (2) Also, some of the data frames don't have Year
 # Need to fix this by joining site-year from weather_final table
 #Soil C Change
soil_C_change <- soil_C_change %>%
  left_join(weather_final %>% select(Site, Year) %>% distinct(), by = "Site")
 # RDM Change
RDM_change <- RDM_change %>%
  left_join(weather_final %>% select(Site, Year) %>% distinct(), by = "Site")

# (3) Need to take the mean of Transect_ID for those that have multiple samples
 # e.g. 3 - 4 soil samples
 # Create a helper function to take the mean of samples for each transect
mean_by_transect <- function(df, keys = c("Year","Site","Treatment","Transect_ID")) {

  # keep only keys that exist in the dataframe
  keys <- keys[keys %in% names(df)]

  df %>%
    group_by(across(all_of(keys))) %>%
    summarise(
      across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
      .groups = "drop"
    )
}

## Transect- Level Response Tables
# Responses: Soil C change (and optional raw C)
soil_C_transect <- soil_C_change %>%
  select(any_of(c(
    "Year","Site","Treatment","Transect_ID",
    "final_Carbon_pct", "final_C_Mg_per_ha",
    "Carbon (%)_Pre-burn", "Carbon (%)_Post-burn",
    "C_Mg_per_ha_Pre-burn","C_Mg_per_ha_Post-burn"
  ))) %>%
  mean_by_transect(keys = c("Year","Site","Treatment","Transect_ID"))

 # Responses: Roots (post-burn only; 2024 burns)
roots_transect <- roots %>%
  select(any_of(c("Year","Site","Treatment","Transect_ID","Roots_g_per_m2"))) %>%
  mean_by_transect(keys = c("Year","Site","Treatment","Transect_ID"))

 # Responses: Aboveground Biomass - RDM change and raw data
RDM_transect <- RDM_change %>%
  select(any_of(c(
    "Year","Site","Treatment","Transect_ID",
    "final_RDM",
    "RDM_g_preburn","RDM_g_postburn"
  ))) %>%
  mean_by_transect(keys = c("Year","Site","Treatment","Transect_ID"))

# Check to make sure they are unique (n=0) ; 1 row per transect
soil_C_transect %>% count(Year, Site, Treatment, Transect_ID) %>% filter(n > 1)
roots_transect  %>% count(Year, Site, Treatment, Transect_ID) %>% filter(n > 1)
RDM_transect    %>% count(Year, Site, Treatment, Transect_ID) %>% filter(n > 1)

# (4) Create a helper function: standardize key columns across dataframes
#    Ensures Site/Treatment/Transect_ID are consistent types
standardize_keys <- function(df) {
  df %>%
    mutate(
      Site = as.character(Site),
      Treatment = if ("Treatment" %in% names(.)) as.character(Treatment) else Treatment,
      Transect_ID = if ("Transect_ID" %in% names(.)) as.character(Transect_ID) else Transect_ID
    )
}

soil_C_change   <- standardize_keys(soil_C_change) #soil carbon
roots           <- standardize_keys(roots) # root biomass
RDM_change      <- standardize_keys(RDM_change) # RDM 
# weather_final   <- standardize_keys(weather_final) # weather predictors
 # No Treatment
# DEM             <- standardize_keys(DEM) # topography
 # No Transect_ID
soil_properties <- standardize_keys(soil_properties) # soil nutrients and texture
CBI             <- standardize_keys(CBI) # Composite Burn Index (CBI)

#

# Build the core "transect master" index
 # Using the union of keys from the response tables
transect_index <- bind_rows(
  soil_C_change %>% select(any_of(c("Year","Site","Treatment","Transect_ID"))),
  roots         %>% select(any_of(c("Year","Site","Treatment","Transect_ID"))),
  RDM_change    %>% select(any_of(c("Year","Site","Treatment","Transect_ID")))
) %>%
  distinct() %>%
  filter(!is.na(Site), !is.na(Treatment), !is.na(Transect_ID))

# Build Transect- Level Predictors Table (each must be 1 row per join key)
# Predictors: Weather (site-level only; cols 4:28)
weather_pred <- weather_final %>%
  select(Site, all_of(names(weather_final)[4:28])) %>%
  distinct(Site, .keep_all = TRUE)

# Predictors: DEM (site-level + treatment; cols 3:5)
DEM_pred <- DEM %>%
  select(Site, Treatment, all_of(names(DEM)[3:5])) %>%
  distinct(Site, Treatment, .keep_all = TRUE)

# 6) Predictors: Soil properties baseline (transect-level; Time == Pre-burn; cols 7:34)
soil_pred_pre <- soil_properties %>%
  filter(Time == "Pre-burn") %>%
  select(Site, Treatment, Transect_ID, all_of(names(soil_properties)[7:34])) %>%
  distinct(Site, Treatment, Transect_ID, .keep_all = TRUE)

# Predictors: CBI (burn-only Treatment; keep other scores)
 # Join by Year+Site+Transect_ID (Treatment is always Burn)
cbi_pred <- CBI %>%
  select(any_of(c(
    "Year","Site","Transect_ID",
    "CBI","CBI_Category",
    "Litter_Score","RDM_Score","Severity_Score"
  ))) %>%
  distinct(Year, Site, Transect_ID, .keep_all = TRUE)

# Final analysis dataframe (predictors + all responses)
 # predictors_master is all of the predictor variables
 # analysis_carbon is all of the predictor variables with carbon response variables
analysis_carbon <- transect_index %>%
  # predictors
  left_join(weather_pred,  by = "Site") %>%
  left_join(DEM_pred,      by = c("Site","Treatment")) %>%
  left_join(soil_pred_pre, by = c("Site","Treatment","Transect_ID")) %>%
  left_join(cbi_pred,      by = c("Year","Site","Transect_ID")) %>%
  # response (mean of samples from each transect)
  left_join(soil_C_transect, by = c("Year","Site","Treatment","Transect_ID")) %>%
  left_join(roots_transect,  by = c("Year","Site","Treatment","Transect_ID")) %>%
  left_join(RDM_transect,    by = c("Year","Site","Treatment","Transect_ID")) %>% 
  left_join(Litter_Pre_burn, by = c("Year","Site","Treatment","Transect_ID"))

# Added in Pre-burn Litter later
analysis_carbon <- analysis_carbon %>%
  left_join(Litter_Pre_burn,
            by = c("Year", "Site", "Treatment", "Transect_ID"))

# Need to get rid of .y and .x
analysis_carbon <- analysis_carbon %>%
  mutate(
    Litter_Pre_burn = coalesce(Litter_Pre_burn.x, Litter_Pre_burn.y)
  ) %>%
  select(-Litter_Pre_burn.x, -Litter_Pre_burn.y)

# Quick QA/QC checks
 # Check how many rows and missingness for key response variables
 # Need 1 row per transect key
dup_check <- analysis_carbon %>%
  count(Year, Site, Treatment, Transect_ID) %>%
  filter(n > 1)

print(dup_check)  # should be 0 rows

qa_summary <- analysis_carbon %>%
  summarise(
    n = n(),
    n_soilC_pct = sum(!is.na(final_Carbon_pct)),
    n_soilC_Mg  = sum(!is.na(final_C_Mg_per_ha)),
    n_roots     = sum(!is.na(Roots_g_per_m2)),
    n_rdm       = sum(!is.na(final_RDM))
  )

print(qa_summary)

# Optional: identify rows missing baseline soil predictors (often a key join issue)
# analysis_carbon %>% filter(is.na(sand) & is.na(silt) & is.na(clay)) %>% select(Year,Site,Treatment,Transect_ID) %>% head(20)

# Look at final merge
head(analysis_carbon)

# Diagnostic for each Site
 # This is the Sample Size Table
analysis_carbon %>%
  summarise(
    soilC = sum(!is.na(final_Carbon_pct)),
    roots = sum(!is.na(Roots_g_per_m2)),
    rdm   = sum(!is.na(final_RDM)),
    .by = Site
  )
## 10 soils for AM, WR22, and CF since it also has Pre- and Post-burn for Controls
## PL roots and RDM have three missing since there are only 2 control transects
## WR5 has a missing soil C since it has only 4 burn transects (1 didn't burn)
## WR5 has one missing RDM (I think it was lost)
```
