---
title: "Data Wrangling"
subtitle: Feb. 24, 2026, Georgia Vasey
format: pdf
gfm: default
pdf: default
execute:
  warning: false
---

## Data Wrangling

I have multiple raw datasets that I would eventually like to merge into one or two dataframes to complete my analyses on.

## 1 - Soil Carbon

First, I need to read in the libraries to help with data wrangling.

```{r warning=FALSE}
## Load packages from library to help with data wrangling
library("tidyverse")
library(here)
```

Now, I'm bringing in the raw soil carbon dataset from Brookside Labs, OH. I want to remove the top seven rows, which is the metadata from Brookside Labs. I need to mutate the column names so that the rest of the raw datasets can match.

```{r}
# Import the data frames for soil carbon
# I collected data in 2023 and 2024 before and after the burns
soil_C_N_2024 <- read_csv(here("data", "soil_C_N_2024.csv"),
                          skip = 7) # Removes top 7 rows with metadata
soil_C_N_2023 <- read_csv(here("data", "soil_C_N_2023.csv"),
                          skip = 7) # Removes top 7 rows with metadata

# Take a look at the data in the first six rows to make sure it looks correct
print(soil_C_N_2023)
```

Now, I want to manipulate the dataframe

```{r}
# See column names for 2024 burns
colnames(soil_C_N_2024)

# Create clean dataframe for soil carbon in 2024 that only includes Carbon
soil_C_2024 <- soil_C_N_2024 %>%
  # Remove everything except for %C
  select(-2, -3, -5, -6) %>%
  # Need to standardize syntax of Sample Location before separating into new columns
  #Create a standardized version of Sample Location so everything uses "_"
  mutate(Sample_Location_std = str_replace_all(`Sample Location`, "-", "_")
  ) %>% 
  # Also need to rename CSR, to remove CSR_SC (I didn't end up sampling at two different sites within CSR)
    mutate(
  `Sample Location` = if_else(
    #The condition: If a Sample Location starts with CSR_SC_
    str_starts(`Sample Location`, "CSR_SC_"), #value_if_true
    # If FALSE, replace with the value_if_true (CSR_SC_) with value_if_false (CSR_)
    str_replace(`Sample Location`, "^CSR_SC_", "CSR_"),
    # If string does not include CSR_SC, return original value unchanged
    `Sample Location`
  )
) %>% 
  # Separate 'Sample Location' into individual columnns
  separate(col = 'Sample Location',
           into = c("Site", "Treatment", "Transect_ID", "Sample", "Time"),
           sep = "_") %>% 
    mutate(Treatment = recode(Treatment, "B" = "Burn", "C" = "Control"),
           Time = recode(Time, "pre" = "Pre-burn","post" = "Post-burn"),
           Sample = as.numeric(Sample),
           'Carbon (%)' = as.numeric(`Carbon   (%)`)) %>% 
  # Get rid of the Carbon column that is character
  select(-6)

# Create clean dataframe for soil carbon in 2023 that only includes Carbon
soil_C_2023 <- soil_C_N_2023 %>%
  # Only include %C
  select(-1, -7, -8, -10, -11) %>% 
  mutate('Carbon (%)' = as.numeric(`Carbon   (%)`)) %>% 
  # Get rid of the Carbon column that is character
  select(-6)

```

Now it's time to merge 2023 and 2024 burns with bulk density, so that I can calculate soil C Mg ha.

```{r}
# Merge soil_C_2023 and soil_C_2024
soil_C <- bind_rows(soil_C_2023, soil_C_2024)

# Import bulk density data frame
bulk_density <- read_csv(here("data", "bulk_density.csv"))

# Only keep bulk density
bulk_density_red <- bulk_density %>% 
  # Keep only Site, Treatment, Name, Bulk Density
  select(3, 4, 5, 12) %>% 
  # Rename bulk density
  rename(bulk_density = `Correct bulk density`) %>%
  # Correct Site names for Wilder Ranch to not include underscore
  mutate(
    Site = str_replace_all(Site, "_", "")
  ) %>% 
  # Add a column called Transect_ID
  mutate(
    Transect_ID = str_extract(Name, "T[1-5]")
  ) %>% 
  #Relocate column Transect_ID
  relocate(Transect_ID, .after = Name)

# Join bulk density into the dataframe
# Also removed MC (Molino Creek), since I didn't send in samples
soil_C_bulk <- soil_C %>%
  left_join(
    bulk_density_red,
    by = c("Site", "Treatment", "Transect_ID")
  ) %>% 
  # Removes Name
  select(-7) 

#Calculate a new column for soil carbon per hectare 
soil_C_bulk <- soil_C_bulk %>% 
  mutate(C_Mg_per_ha = `Carbon (%)`*bulk_density*10 )
```
